---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
rmarkdown::render(

## Loading and preprocessing the data

###### Read in the raw data set

```{r}
step_data_raw <- read.csv("activity.csv")
step_data_no_na <- na.omit(step_data_raw)
```

## What is mean total number of steps taken per day?

###### Use dpylr package to manipulate data

```{r}
library(dplyr)
step_data_sum <- summarise(group_by(step_data_no_na, date), total_steps = sum(steps))
```

###### Mean and Median for the data excluding NA values

```{r}
mean(step_data_sum$total_steps) 
median(step_data_sum$total_steps) 
```

```{r}
hist(step_data_sum$total_steps, main = "Total Steps per Day", xlab = "Sum of Steps Per Day", ylab = "Count of Days")
```


## What is the average daily activity pattern?


```{r}
step_data_means <- summarise(group_by(step_data_no_na, interval), average_steps = mean(steps))
```

```{r}
plot(step_data_means$interval, step_data_means$average_steps, xlab = "Time Interval", ylab = "Average Steps Per Inteval", type = "l")
```

###### Max average steps for daily intervals

```{r}
top_6_average_steps <- head(step_data_means[order(-step_data_means$average_steps, step_data_means$interval), ])
max_average_steps <- top_6_average_steps[1,]
max_average_steps
```

## Imputing missing values

###### I am setting the missing values by taking average total number of steps per day (10766.19) and dividing by the number of intervals (288) in a day
###### average number of steps per interval = 10766.19 / 288 --> 37.4

```{r}
step_data_replace <- step_data_raw %>% mutate(replace(step_data_raw$steps, which(is.na(step_data_raw$steps)),37.4))
colnames(step_data_replace) <- c("steps_na", "date", "interval", "steps")
step_data_sum_2 <- summarise(group_by(step_data_replace, date), total_steps = sum(steps))
mean(step_data_sum_2$total_steps)
median(step_data_sum_2$total_steps)
```

###### Difference in mean comparing NA data set and replaced NA data set --> 0.66 steps
###### Difference in median comparing NA data set and replaced NA data set --> 6.2 steps

```{r}
hist(step_data_sum_2$total_steps, main = "Total Steps per Day (with replaced values)", xlab = "Sum of Steps Per Day", ylab = "Count of Days")
```

## Are there differences in activity patterns between weekdays and weekends?

###### create new data set for weekday analysis by creating an additional column to identify weekday vs weekend

```{r}
library(chron)
week_col <- is.weekend(as.Date(step_data_replace$date))
step_data_week <- cbind(step_data_replace, week_col)
colnames(step_data_week) <- c("steps_na", "date", "interval", "steps", "is_weekend?")
```

###### summarise data set then plot comparing weekday vs weekend trends

```{r}
step_data_week_inter <- summarise(group_by(step_data_week, interval, `is_weekend?`), average_steps = mean(steps))
library(ggplot2)
qplot(step_data_week_inter$interval, step_data_week_inter$average_steps, group = `is_weekend?`, data = step_data_week_inter, geom = "line", col = `is_weekend?`, xlab = "Time Interval", ylab = "Average Steps", main = "Average Steps Weekend Vs WeekDay") + scale_color_manual(labels = c("Weekday", "Weekend"), values = c("red", "blue"))
```
)
```
